# Case study

Typical actions you perform on your data:

- Import
- Explore
- Model
- Visualize
- Report

```{r}
# Prepare R session
setwd("~/courses/talk-r-movies")
```

## Import

- R can deal with many sources
- Online and offline
- Dedicated application-specific packages (DBs, excel, text files)

```{r}
# read one part from database
library(RSQLite)
con <- dbConnect(SQLite(), "omdb.db")
dbListTables(con)
omdb <- dbReadTable(con, "omdb")

# read.csv is a built-in function
tomatoes <- read.csv("rotten_tomatoes.csv", header = TRUE)
```

## Explore

- Get a first feel for your data: `summary()`, `head()`, `tail()`, `names()`.

```{r}
# explore omdb
names(omdb)
head(omdb)
summary(omdb)

# explore tomatoes
names(tomatoes)
tail(tomatoes)
summary(tomatoes)
```

- Merge `tomatoes` with `omdb` to get one data set

```{r}
# Do the merge
movies <- merge(omdb, tomatoes, by="ID")

# Check out the result
names(movies)
head(movies)
summary(movies)
```

- There is still some cleaning needed!

```{r}
# let's work with imdbRating and userRating
names(movies)[names(movies) == "userRating"] <- "rottenRating"
names(movies)[names(movies) == "userReviews"] <- "rottenVotes"

# Remove NA rating info, before it messes up our analyses
row_misses_info <-  is.na(movies$imdbVotes) | is.na(movies$rottenVotes) | is.na(movies$imdbRating) | is.na(movies$rottenRating)
summary(row_misses_info)
movies <- movies[!row_misses_info, ]

# Remove movies that haven't got 10000 votes for both imdb and rottentomatoes
movies <- movies[movies$imdbVotes > 10000 & movies$rottenVotes > 10000, ]

# Keep only columns that we will be using later on
cols_to_keep <- c("Title", "Year", "Runtime", "imdbRating", "imdbVotes", "rottenRating", "rottenVotes")
movies <- movies[cols_to_keep]

# Check out the result
summary(movies)    # looks good!
```

Let's make one overall rating for movies (columns can be added using `$` notation)

```{r}
movies_dplyr <- movies

# make one general rating
movies$Votes <- movies$imdbVotes + movies$rottenVotes
# weighted average
movies$Rating <- (movies$imdbVotes*movies$imdbRating + movies$rottenVotes*2*movies$rottenRating)/(movies$Votes)


# dplyr alternative:
library(dplyr)
movies_dplyr <- movies_dplyr %>% 
  mutate(Votes = imdbVotes + rottenVotes, 
         Rating = (imdbVotes*imdbRating + rottenVotes*2*rottenRating)/Votes)
```

Let's do some first exploratory queries:

```{r}
# Top rated movies
head(movies[order(-movies$Rating), c("Title", "Rating")],15)

# dplyr alternative
movies %>% select(Title, Rating) %>% arrange(desc(Rating)) %>% head(15)

# Most voted movies
head(movies[order(-movies$Votes), c("Title", "Votes")], 15)

# dplyr alternative
movies %>% select(Title, Votes) %>% arrange(desc(Votes)) %>% head(15)
```

## Model

Is there a correlation between a movie's runtime and its rating?

```{r}
cor(movies$Rating, movies$Runtime)
```

```{r}
plot(movies$Runtime, movies$Rating)
lmfit <- lm(movies$Rating ~ movies$Runtime)
lmfit
summary(lmfit)
```

```{r}
lmfit2 <- lm(movies$Rating ~ movies$Runtime + movies$Year)
lmfit2
plot(lmfit2)
```

## Visualize

```{r}
hist(movies$Rating, breaks = 25)
boxplot(movies$Rating)
```

Compare rotten ratings to imdb ratings

```{r}
hist(movies$imdbRating, breaks = 25)
hist(movies$rottenRating*2, breaks = 25)
```

```{r}
par(mfrow = c(1, 2))
boxplot(movies$rottenRating*2)
boxplot(movies$imdbRating)
par(mfrow=c(1,1))
```

Connection between runtime and rating that is not linear?

```{r}
library(ggplot2)
qplot(Runtime, Rating, data = movies) +  stat_smooth()

library(ggvis)
movies %>% ggvis(~Runtime, ~Rating) %>% layer_points() %>% layer_smooths()
```

## Report

- Reproducible reporting: R Markdown
- R code chunks inside a document
- Uses Markdown syntax

## Powerful packages

- dplyr (easy) & data.table (fast, "i, j, by")
- ggvis (powerful grammar)
- Free DataCamp Trial: www.datacamp.com/demo
