# Case study

Typical actions you perform on your data:

- Import
- Clean
- Explore & Visualize
- Model
- Report

Goal of this case study: investigate movie ratings from IMdB and Rotten Tomatoes.

```{r}
# Prepare R session
# setwd("cs_movies")
rm(list = ls())
```

## Import

- R can deal with many sources, online and offline
- Dedicated application-specific packages (DBs, excel, text files)

We read the IMdB data from an SQLite database:

```{r, echo = FALSE}
# install.packages("RSQLite")
library(RSQLite)
con <- dbConnect(SQLite(), "omdb.db")
dbListTables(con)
omdb <- dbReadTable(con, "omdb")
```

The rotten tomatoes data comes as a CSV, that we can read in with `read.csv()`.

```{r, echo = FALSE}
tomatoes <- read.csv("rotten_tomatoes.csv")
```

## Explore

Get a first feel for your data: `summary()`, `head()`, `tail()`, `names()`.

```{r}
# explore omdb
names(omdb)
head(omdb)
summary(omdb)

# explore tomatoes
names(tomatoes)
tail(tomatoes)
summary(tomatoes)
```

Use the `merge()` function to merge `tomatoes` and `omdb`: one data set. Looking at the names, they have the `ID` column in common.

```{r}
movies <- merge(omdb, tomatoes, by = "ID")

# Check out the result
names(movies)
head(movies)
summary(movies)
```

## Clean

There is still some cleaning needed!

1. Rename some columns
2. Remove observation with NA rating info
3. Remove observations with less than 10000 votes
4. Select columns of interest

```{r}
# install.packages("dplyr")
library(dplyr)
movies_clean <- movies %>%
  rename(rottenRating = userRating, rottenVotes = userReviews) %>%
  filter(!is.na(imdbVotes), !is.na(rottenVotes), !is.na(imdbRating), !is.na(rottenRating)) %>%
  filter(imdbVotes > 10000, rottenVotes > 10000) %>%
  select(Title, Year, Runtime, imdbRating, imdbVotes, rottenRating, rottenVotes)
```

If we have a look at the summary, it looks much better now!

```{r}
summary(movies_clean)
```

Let's also make one overall rating for movies (columns can be added using `$` notation): count the number of votes, create a weighted average of `rottenVotes` and `rottenRating`. IMdB rating are /10, Rotten Tomatoes votes are /5, take this into account!

```{r}
movies_overall <- movies_clean %>% 
  mutate(Votes = imdbVotes + rottenVotes, 
         Rating = (imdbVotes * imdbRating + rottenVotes * 2 * rottenRating) / Votes)
```


## Explore

Let's do some first exploratory queries.

List of top-rated movies:

```{r}
movies_overall %>% 
  select(Title, Rating) %>% 
  arrange(desc(Rating)) %>% 
  head(15)
```

List of movies that got the most votes:

```{r}
movies_overall %>% 
  select(Title, Votes) %>% 
  arrange(desc(Votes)) %>% 
  head(15)
```

## Visualize

Plot the ratings against the year:

```{r}
library(ggvis)
movies_overall %>% 
  ggvis(~Year, ~Rating) %>% 
  layer_points()
```

How are ratings distributed?

```{r}
movies_overall %>% 
  ggvis(~Rating) %>% 
  layer_histograms(width = input_slider(0.1, 2, step = 0.1, value = 0.5, label = "Change binwidth"))
```

Compare rotten ratings to imdb ratings

```{r}
par(mfrow = c(2, 1))
hist(movies_overall$imdbRating, breaks = 25, ylim = c(0, 1000))
hist(movies_overall$rottenRating*2, breaks = 25, ylim = c(0, 1000))
par(mfrow = c(1,1))
```

## Model

Is there a correlation between a movie's runtime and its rating?

```{r}
cor(movies_overall$Rating, movies_overall$Runtime)
plot(movies_overall$Rating, movies_overall$Runtime)
```

Let's use the `ggvis` package to plot a non-linear model to explain Rating based on Runtime:

```{r}
library(ggvis)
movies_overall %>% 
  ggvis(~Runtime, ~Rating) %>% 
  layer_points(opacity := 0.1) %>% 
  layer_smooths(se = TRUE)
```

## Report

- This file is an example of reproducible reporting!
- R Markdown: R code and normal markdown text interwoven
- Change R code: report changes accordingly
- Simply hit Knit HTML
