# Fundamentals

- Explore the console: simple calculations
- Variable assignment: `<-`
- Different data types: numeric, character, logical
- As you execute R commands, you accumulate a workspace.
- R scripts.
- Lines of R code, executed one after the other

```
score <- 12
course <- "algebra"
passed <- TRUE
ls()
```

# Data structures

## Vectors

- 1 Dimensional
- Single type
- `c()`
- Subsetting with `[ ]`
- Single numbers are also vectors, of length 1!

```{r}
# creation
scores <- c(12, 17, 19, 9)
scores
courses <- c("algebra", "physics", "management", "religion")
courses

# single type!
c("algebra", 12) # coercion

# naming
scores <- c(algebra = 12, physics = 17, management = 19, religion = 9)
scores

# subsetting
scores[2]
scores["physics"]
```

## Lists

- 1 Dimensional
- Different types: numerics, logicals, vectors, lists ...
- `list()`
- Subsetting with `[[ ]]`

```{r}
summ <- list("algebra", 12, TRUE, "physics", 17, TRUE, "management", 19, TRUE, "religion", 9, FALSE)
summ
str(summ)

summ2 <- list(list("algebra", 12, TRUE), 
              list("physics", 17, TRUE), 
              list("management", 19, TRUE), 
              list("religion", 9, FALSE))
str(summ2)


summ3 <- list(list(course = "algebra", score = 12, passed = TRUE),
              list(course = "physics", score = 17, passed = TRUE),
              list(course = "management", score = 19, passed = TRUE),
              list(course = "relgion", score = 9, passed = FALSE))
str(summ3)

summ3[[1]]
summ3[[1]][[1]]
summ3[[1]]$course

summ3[[3]][[2]]
summ3[[3]]$score
```

## Data Frames

- 2 Dimensional
- Heterogeneous content
- Under the hood: a named list with same-length vectors
- Typical structure for data sets: rows are observations, columns are variables.
- `data.frame()` function
- Different ways of subsetting

```{r, error = TRUE}
options(stringsAsFactors=FALSE)

courses <- c("algebra", "physics", "management", "religion")
scores <- c(12, 17, 19, 9)
passed <- c(TRUE, TRUE, TRUE, FALSE)

results <- data.frame(courses, scores, passed)
results

results[2, 1] # 2nd row, 1st column
results[2, ] # entire 2nd row
results[, 1] # entire 1st column
results$courses

# data frame is a list!
str(results)
```

# Functions and Packages

## Functions

- Tons of them available in R.
- Solve particular, well-defined problem.
- You already used them! `c()`, `list()`, etc.
- Black Box Principle
- Can also write your own.
    
```
# What is your average score?
results
results$scores
mean(results$scores) # mean is a base R function
```

## R Packages

R Packages: sets of R functions, data and compiled code to help solve a typical problem. Packages for visualization, high frequency trading, advanced statistical analyses, machine learning. Most of them are on [CRAN](http://cran.r-project.org/), the _Comprehensive R Archive Network_.

```{r error = TRUE}
datatable(results) # function the DT package

install.packages("DT")
library(DT)
datatable(results)
```

## Let's practice!

- Create account on https://www.datacamp.com/
- Head over to https://www.datacamp.com/courses/r-workshop and start the mini-course

# Other programming concepts (extra)

## Control Structures

- use results of comparators to customize behavior

```{r}
temp_celsius <- 68
if(temp_celsius <= 0) {
  print("Brrrr!")
} else {
  print("It's not freezing.")
}

color <- "red"
if(color == "green") {
  print("You love nature!")
} else if(color == "red") {
  print("You like campfires!")
} else {
  print("I can't really tell what you like..")
}
```
    
## Loops
    
- repeated execution of code chunks

```{r}
for(i in 1:5) {
  print(i)
}

hobbies <- c("cycling", "movies", "data science")
for(h in hobbies) {
  print(h)
}
```

