---
title: "Plotting unemployment in Germany"
author: "Filip Schouwenaars"
date: "April 26, 2016"
output: html_document
---

## Preparation

```{r}
# If not installed: uncomment line below
# install.packages(c("rgdal", "ggplot2", "ggmap", "ggthemes", "viridis", "rvest"))
```

```{r}
# Change to path of cs_maps/
setwd("~/workspace/r-workshop/cs_maps/")
```

## Shape files

Shapefiles can be used to describe points, polylines or polygons. A single shapefile actually consists of several files, each describing a specific aspect of the overall geometry. 

- `.shp` describes the shape, the feature geometry itself.
- `.shx` represents the shape index, a positional index.
- `.dbf` contains the attributes for each shape arranged in columns.

The prefix name of these files must be consistent and they must be kept in the same directory. Here we have data for Germany, all starting with `DEU`. The suffix specifies the level of organization: 

- `DEU_adm0` is the administrative (political) boundaries of the entire country
- `DEU_adm1` is the administrative boundaries for each of the 16 states

We will use `rgdal::readOGR()` to import the `DEU_adm1` data.

Other packages: `rgeos`, `maptools`, `sp`

```{r}
library(rgdal) # for getting shapefiles
bundes_raw <- readOGR(dsn = "germany", layer = "DEU_adm1")
class(bundes_raw)
plot(bundes_raw)
```

Let's try to plot this somewhat nicer and more extensible, with `ggplot2`. We will use `fortify` to convert the `SpatialPolygonsDataFrame` to a more suitable format.

```{r}
library(ggplot2)
library(ggmap) # for themes
library(ggthemes) # for themes
library(viridis) # for a nice color theme

fortify # S3 function (multiple dispatch) 
ggplot2:::fortify.SpatialPolygonsDataFrame
ggplot2:::fortify.Polygon
length(bundes_raw@polygons[[1]]@Polygons)

# Convert SpatialPolygonsDataFrame to data.frame
bundes <- fortify(bundes_raw)

# Display structure
str(bundes)

# Stupid plot
plot(bundes$long, bundes$lat, pch = ".")

# Nice plot
ggplot(bundes, aes(x = long, y = lat, group = group)) +
    geom_polygon(fill = "#f4cb42", col = "#44434c") +
    coord_map() + 
    theme_nothing()
```

Suppose we now have additional data on unemployment in Germany's different Bundes.

```{r}
unemp <- read.delim("germany_unemployment.txt")
str(unemp)

# Add names to bundes so you can merge
bundes$state <- factor(as.numeric(bundes$id))
levels(bundes$state) <- bundes_raw$NAME_1

# Merge bundes and unemp, on state
bundes_unemp <- merge(bundes, unemp)

# Create plot, colored according to unemployment
ggplot(bundes_unemp, aes(x = long, y = lat, group = group, fill = unemployment)) +
  geom_polygon() +
  coord_map() +
  theme_map() + 
  scale_fill_viridis()
```

We can take it one stap further by superimposing german cities and size according to population:

```{r}
# Get coordinates of major cities, with population estimates
cities <- read.csv("german_cities.csv")

# Plot different layers
ggplot(bundes_unemp) +
  geom_polygon(aes(x = long, y = lat, group = group, fill = unemployment)) +
  geom_point(data = cities, aes(x = lon, y = lat, size = Population),
             col = "red", shape = 16, alpha = 0.7) +
  scale_size(range = c(0,15)) +
  coord_map() +
  theme_map() +
  scale_fill_viridis()
```

No clear connection...

## Extra: scrape german cities info

How the german cities data frame came about:

```{r, eval = FALSE}
write("
// scrape_techstars.js

var webPage = require('webpage');
var page = webPage.create();

var fs = require('fs');
var path = 'germancities.html'

page.open('http://www.citymayors.com/gratis/german_topcities.html', function (status) {
  var content = page.content;
  fs.write(path,content,'w')
  phantom.exit();
});
", file = "scrape_web.js")

# First get phantomjs from http://phantomjs.org/
system("./phantomjs scrape_web.js")

library(rvest)
lines <- read_html("germancities.html", encoding = "UTF-8") %>%
  html_nodes("table tbody tr td table tbody tr") %>% html_text()

# drop last
splitup <- strsplit(lines[-length(lines)], "\t|\n")
clean <- lapply(splitup, function(x) gsub("^\\s+|\\s+$", "", x[x != ""]))
df <- as.data.frame(do.call(rbind, clean[-1]), stringsAsFactors = FALSE)
names(df) <- clean[[1]]
df$Rank <- as.numeric(df$Rank)
df$Population <- as.numeric(gsub(",", "", df$Population))
df$Rank <- NULL
xy <- ggmap::geocode(location = df$City)
xy$City <- df$City
ultra_df <- merge(df, xy)
write.csv(ultra_df, file = "german_cities.csv")
```

## About this report

This is a R Markdown document, which is fully reproducible. Simply hit 'Knit HTML' inside RStudio, and a html document will be built. Very easily, you can also turn it into \LaTeX backed PDF documents. You can update the options inside the R chunks to display or hide R code, output, plots, results, etc. You can use `kable()`, which is baked into R Markdown, or packages like `pander` to create good-looking tables.

## References

These scripts are based on work that Rick Scavetta did for DataCamp, in his [third course on ggplot2](https://www.datacamp.com/courses/862).
