---
title: "Plotting unemployment in Germany"
author: "Filip Schouwenaars"
date: "April 26, 2016"
output: html_document
---

```{r}
setwd("~/workspace/r-workshop/cs_maps/")
```

Shapefiles can be used to describe points, polylines or polygons - here you'll focus on polygons for drawing maps.

A single shapefile actually consists of several files, each describing a specific aspect of the overall geometry. The three core file types are:

- `.shp`: the shape, the feature geometry itself.
- `.shx`: the shape index, a positional index.
- `.dbf`: the attribute, attributes for each shape arranged in columns.

The prefix name of these files must be consistent and they must be kept in the same directory. The files you'll use are for Germany, and all begin with `DEU`. The suffix specifies the level of organization: 

- `DEU_adm0` is the administrative (political) boundaries of the entire country
- `DEU_adm1` is the administrative boundaries for each of the 16 states

```{r}
library(rgdal) # for getting shapefiles
library(ggplot2) # for 'fortifying'
library(ggmap); library(ggthemes) # for themes

# country level
germany_raw <- readOGR(dsn = "germany", layer = "DEU_adm0")
class(germany_raw)
fortify # S3 function
ggplot2:::fortify.SpatialPolygonsDataFrame
germany <- ggplot2:::fortify.SpatialPolygonsDataFrame(germany_raw)
str(germany)
ggplot(germany, aes(x = long, y = lat, group = group)) +
    geom_polygon(fill = "blue", col = "white") +
    coord_map() + 
    theme_nothing()

# state level (bundes)
bundes_raw <- readOGR(dsn = "germany", layer = "DEU_adm1")
class(bundes_raw)
bundes <- fortify(bundes_raw)
str(bundes)
ggplot(bundes, aes(x = long, y = lat, group = group)) +
    geom_polygon(fill = "blue", col = "white") +
    coord_map() + 
    theme_nothing()

## Color according to unemployment

# Add names again
bundes$state <- factor(as.numeric(bundes$id))
levels(bundes$state) <- bundes_raw$NAME_1

# Import unemployment data
unemp <- read.delim("germany_unemployment.txt")

# merge bundes and unemp
bundes_unemp <- merge(bundes, unemp)

# Choropleth
ggplot(bundes_unemp, aes(x = long, y = lat, group = group, fill = unemployment)) +
  geom_polygon() +
  coord_map() +
  theme_map()

# Get coordinates of major cities, with population estimates
cities <- read.csv("german_cities.csv")
ggplot(bundes_unemp, aes(x = long, y = lat, fill = unemployment)) +
  geom_polygon(aes(group = group)) +
  geom_point(data = cities, aes(size = Pop_est),
             col = "red", shape = 16, alpha = 0.6) +
  coord_map() +
  theme_map()

```

How the german cities data frame came about:

```{r}
write("
// scrape_techstars.js

var webPage = require('webpage');
var page = webPage.create();

var fs = require('fs');
var path = 'germancities.html'

page.open('http://www.citymayors.com/gratis/german_topcities.html', function (status) {
  var content = page.content;
  fs.write(path,content,'w')
  phantom.exit();
});
", file = "scrape_web.js")
system("./phantomjs scrape_web.js")

lines <- read_html("germancities.html", encoding = ) %>% guess_encoding()
  html_nodes("table tbody tr td table tbody tr") %>% html_text()

# drop last
splitup <- strsplit(lines[-length(lines)], "\t|\n")
clean <- lapply(splitup, function(x) gsub("^\\s+|\\s+$", "", x[x != ""]))
df <- as.data.frame(do.call(rbind, clean[-1]), stringsAsFactors = FALSE)
names(df) <- clean[[1]]
df$Rank <- as.numeric(df$Rank)
df$Population <- as.numeric(gsub(",", "", df$Population))
class(batches)
df$Rank <- NULL
xy <- ggmap::geocode(location = df$City)
xy$location <- df$City
ultra_df <- merge(df, xy)
write.csv(ultra_df, file = "german_cities.csv")
```

## About this report

This is a R Markdown document, which is fully reproducible. Simply hit 'Knit HTML' inside RStudio, and a html document will be built. Very easily, you can also turn it into \LaTeX backed PDF documents. You can update the options inside the R chunks to display or hide R code, output, plots, results, etc. You can use `kable()`, which is baked into R Markdown, or packages like `pander` to create good-looking tables.

## References

These scripts are based on work that Rick Scavetta did for DataCamp, in his [third course on ggplot2](https://www.datacamp.com/courses/862)
