---
title: "Analyzing library data"
author: "Filip Schouwenaars"
date: "April 26, 2016"
output: pdf_document
---

Let's discuss the occupation of Leuven's libraries.

## Importing the data

```{r}
# csv_path <- "https://s3.amazonaws.com/documents.datacamp.com/quivr/library_data.csv"
csv_path <- "~/Downloads/library_data.csv"

library(readr)
lib <- read_csv(csv_path)

library(dplyr)
glimpse(lib)
```

## Clean the data

It seems that there is one column that doesn't mean anything; let's do some cleaning to remove it. Also, we'll convert the `date_created` column to an actual date.

```{r}
library(dplyr)
lib_clean <- lib %>% 
  select(-1) %>%
  mutate(date = as.Date(date_created))

head(lib_clean, 10)

# library(knitr)
# kable(head(lib_clean, 10))
```

## Explore and visualize the data

What was the highest occupancy for each library?

```{r}
lib_clean %>%
  group_by(name) %>%
  summarise(max_occ = max(occupancy)) %>%
  arrange(desc(max_occ))

# %>% kable
```

It could be interesting to see the daily average occupation of the AGORA learning center over time. First let's create a data frame with the correct information:

```{r}
agora_daily <- lib_clean %>%
  filter(library_id == 1) %>%
  group_by(date) %>%
  summarise(avg_occ = mean(occupancy))
```

Now we can use `ggplot2`'s grammar of graphics to create a pretty graph

```{r}
library(ggplot2)
ggplot(agora_daily, aes(x = date, y = avg_occ)) + 
  geom_line(col = "red") + 
  ggtitle("Agora occupancy over time")
```

Clear peaks towards the exam periods, and drops during the holidays.

## Model the data

We could do time series modeling to predict the occupancy of the libraries based on historical data. Skipped for now ;-)

## Report the data

This is a R Markdown document, which is fully reproducible. Simply hit 'Knit HTML' inside RStudio, and a html document will be built. Very easily, you can also turn it into \LaTeX backed PDF documents. You can update the options inside the R chunks to display or hide R code, output, plots, results, etc. You can use `kable()`, which is baked into R Markdown, or packages like `pander` to create nice looking tables.
